---
title: "Untitled"
author: "Brendan F. Miller"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BCL layers 1-4

# Intersect OD genes

```{r}

# save(bcl_countsClean,
#       bcl_CPM, # 12821 filteres genes and 1029 spots after cleaning and CPM normalized
#       bclCorpus, # corpus of the 152 genes and 1029 spots (2 dropped)
#       bcl_k12, # LDA trained on OD gene intersection (152 genes between layers 1-4 with alpha = 0.2)
#      file = "/Users/brendan/Desktop/PostDoc/work/STDeconvolve/repos/STdeconvolve/data/bcl_fitted_models.RData")

load("/Users/brendan/Desktop/PostDoc/work/STDeconvolve/repos/STdeconvolve/0draft_data/bcl_fitted_models.RData")

```

```{r}

fig_path <- "/Users/brendan/Desktop/PostDoc/work/STDeconvolve/repos/STdeconvolve/figures/"
draft_fig_path <- "/Users/brendan/Desktop/PostDoc/work/STDeconvolve/repos/STdeconvolve/0draft_figures/"

```

```{r}

data(BCL)
# 4 layers, 1031 total spots.

bcl_pos <- BCL$pos[,1:2]
bcl_slice <- BCL$pos[,3]
names(bcl_slice) <- rownames(bcl_pos)
length(bcl_slice)

# filter spots in bcl_slice based on those kept in the corpus
bcl_slice_filt <- bcl_slice[names(bcl_slice) %in% rownames(bclCorpus$pos)]
length(bcl_slice_filt)

bclCorpus$slice <- bcl_slice_filt

# clean
# bcl_countsClean <- MERINGUE::cleanCounts(counts = BCL$counts, # genes x spots mtx
#                                         min.reads = 10,
#                                         min.lib.size = 10,
#                                         verbose=TRUE)
# 
# bcl_CPM <- MERINGUE::normalizeCounts(counts = bcl_countsClean,
#                                           verbose=TRUE, log=FALSE)

```

The shared OD genes across the 4 layers:

```{r}

# list of either the mob data sets or paths to the data for inputs into `preprocess`

bcl_paths <- list()
bcl_paths[[1]] <- as.matrix(t(BCL$counts[,names(bcl_slice[bcl_slice == 1])]))
bcl_paths[[2]] <- as.matrix(t(BCL$counts[,names(bcl_slice[bcl_slice == 2])]))
bcl_paths[[3]] <- as.matrix(t(BCL$counts[,names(bcl_slice[bcl_slice == 3])]))
bcl_paths[[4]] <- as.matrix(t(BCL$counts[,names(bcl_slice[bcl_slice == 4])]))

```

```{r}

# collect OD genes for each slice
bcl_ODgenes <- lapply(bcl_paths, function(p) {
  dat <- preprocess(p,
                   alignFile = NA,
                   extractPos = FALSE,
                   selected.genes = NA,
                   nTopGenes = NA, # top expressed genes to remove
                   genes.to.remove = NA,
                   perc.spots = NA,
                   min.reads = 10,
                   min.lib.size = 10,
                   ODgenes = TRUE, # select the overdispersed genes
                   od.genes.alpha = 0.2, # more flexible because taking shared ODs will reduce drastically
                   gam.k = 5,
                   nTopOD = NA) # option to keep just the top n OD genes
  colnames(dat$corpus)
})

# get intersection of the OD genes
sharedBCLGenes <- Reduce(intersect, bcl_ODgenes)
length(sharedBCLGenes)

```

## Txn clustering

In the original paper, 3 clusters ID:
- ductal carcinoma in situ
- invasive ductal carcinoma
- non-malignant

```{r}

set.seed(888)

# use the CPM adjusted counts

pcs.info <- prcomp(t(log10(as.matrix(bcl_CPM[sharedBCLGenes,rownames(bclCorpus$pos)])+1)), center=TRUE)
plot(pcs.info$sdev[1:30])
nPcs <- 5
pcs <- pcs.info$x[,1:nPcs]
# 2D embedding by tSNE
emb <- Rtsne::Rtsne(pcs,
                    is_distance=FALSE,
                    perplexity=30,
                    num_threads=1,
                    verbose=FALSE)$Y
rownames(emb) <- rownames(pcs)
# Graph-based cluster detection
k <- 350
bcl_com <- MERINGUE::getClusters(pcs, k, method=igraph::cluster_louvain)


# pdf(file = paste0(draft_fig_path, "Fig_3_txn-clustering-shared-ODs.pdf"),   # The directory you want to save the file in
  # width = 6, # The width of the plot in inches
  # height = 5) # The height of the plot in inches
par(mfrow=c(2,2), mar=rep(1,4))
# top row
# left: all slice spots tSNE embedding and labeled by cluster assignment
MUDAN::plotEmbedding(emb, groups=bcl_com,
                     show.legend=TRUE, xlab=NA, ylab=NA,
                     verbose=FALSE)
# right: all slice spots tSNE embedding and labeled by tissuse slice assignment
MUDAN::plotEmbedding(emb, groups=bclCorpus$slice,
                     show.legend=TRUE, xlab=NA, ylab=NA,
                     verbose=FALSE)
# bottom row
# left: spots labeld by cluster assignment and overlayed on each slice.
# Same clusters seem to be at similar spot locations on slices. Nice to see
MUDAN::plotEmbedding(bclCorpus$pos, groups=bcl_com,
                     cex=1, xlab=NA, ylab=NA,
                     verbose=FALSE)
# right: spots colored by slice and slices overlayed
MUDAN::plotEmbedding(bclCorpus$pos, groups=bclCorpus$slice,
                     cex=1, xlab=NA, ylab=NA,
                     verbose=FALSE)
# dev.off()

# capture the txn cluster assignment of the spots
bclCorpus$txn <- bcl_com

```

```{r}

bclTxnClustCols <- c('1' = lighten('red', factor = 0.4),
                     '2' = lighten('green', factor = 0.4),
                     '3' = lighten('blue', factor = 0.4))

```

```{r}

# proxy theta for the txn clusters
bclProxyTheta <- model.matrix(~ 0 + bclCorpus$txn)
rownames(bclProxyTheta) <- names(bclCorpus$txn)
# fix names
colnames(bclProxyTheta) <- c("1", "2", "3")
bclProxyTheta <- as.data.frame.matrix(bclProxyTheta)
dim(bclProxyTheta)

```

```{r}

for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  # m <- bcl_shared_k3$theta[slice,]
  p <- bclCorpus$pos[slice,]
  # sliceTxnClusters <- bclCorpus$txn[slice]
  m <- as.matrix(bclProxyTheta[slice,])

  plt <- vizAllTopics(theta = m,
                       pos = p,
                       topicOrder=colnames(m),
                       topicCols=as.vector(bclTxnClustCols[colnames(m)]),
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s))
  print(plt)
  
}

```

# Union of OD genes

152 genes useful but maybe capture more variation if keep OD genes regardless of the slice (take union)

```{r}

# lets also make sure we are using genes that are present across all slices
# because the final corpus will be all the spots across all the slices
# and each one will also need to contain the genes in the union
all_genes <- lapply(bcl_paths, function(p){
  print(dim(p))
  genes <- colnames(p)
  print(length(genes))
  genes
})
all_genes <- Reduce(intersect, all_genes)
length(all_genes)

```
seems like each slice has the same genes anyways

```{r}

# find shared set of OD genes across sections to try and identify common transcriptional patterns

bcl_ODgenes <- lapply(bcl_paths, function(p) {
  dat <- preprocess(p,
                   alignFile = NA,
                   extractPos = FALSE,
                   selected.genes = all_genes, # use the set of shared genes across all slices
                   nTopGenes = 5,
                   genes.to.remove = NA,
                   perc.spots = 0.95,
                   min.reads = 10,
                   min.lib.size = 10,
                   min.detected = 1,
                   ODgenes = TRUE, # select the overdispersed genes
                   od.genes.alpha = 0.05, # use a smaller alpha to use the union of the top OD genes in each section
                   gam.k = 5,
                   nTopOD = 100) # it turns out the final list is too large for the topic modeling. So select just the top OD genes in each slice
  colnames(dat$corpus)
})

# because very few genes are OD across all sections, instead take the union and consider genes
# that are OD in any slice
unionBCLGenes <- Reduce(union, bcl_ODgenes)
length(unionBCLGenes)

```

Build the corpus with the union set of OD genes. Here its 372. 

```{r}

bclCorpusUnion <- preprocess(t(as.matrix(BCL$counts)),
                       alignFile = NA,
                       extractPos = FALSE,
                       selected.genes = unionBCLGenes,
                       nTopGenes = NA,
                       genes.to.remove = NA,
                       perc.spots = NA,
                       min.reads = 0, 
                       min.lib.size = 0, # because using specific set of genes, remove spots with 0 counts but keep others
                       ODgenes = FALSE,
                       od.genes.alpha = 0.05, # ODgenes ignored; so is this
                       gam.k = 5) # ODgenes ignored; so is this

# positions already included in `mOB$pos` but make sure only use the filtered positions in the corpus
bclCorpusUnion$pos <- bcl_pos[rownames(bclCorpusUnion$corpus), ]

# add in slice info for the spots that were kept after filtering with the new gene set.
# With the new genes, some different spots may not have any gene counts and will be empty and should be removed
bcl_slice_filt <- bcl_slice[names(bcl_slice) %in% rownames(bclCorpusUnion$pos)]
length(bcl_slice_filt)

bclCorpusUnion$slice <- bcl_slice_filt

```

## Txn clustering

txn clustering of the spots using the union set of OD genes

```{r}

set.seed(888)

# use the CPM adjusted counts

pcs.info <- prcomp(t(log10(as.matrix(bcl_CPM[unionBCLGenes,rownames(bclCorpusUnion$pos)])+1)), center=TRUE)
plot(pcs.info$sdev[1:30])
nPcs <- 5
pcs <- pcs.info$x[,1:nPcs]
# 2D embedding by tSNE
emb <- Rtsne::Rtsne(pcs,
                    is_distance=FALSE,
                    perplexity=30,
                    num_threads=1,
                    verbose=FALSE)$Y
rownames(emb) <- rownames(pcs)
# Graph-based cluster detection
k <- 350
bcl_com <- MERINGUE::getClusters(pcs, k, method=igraph::cluster_louvain)

bcl_colors <- bcl_com
levels(bcl_colors)[levels(bcl_colors) == "1"] <- "green" #"#66FF66"
levels(bcl_colors)[levels(bcl_colors) == "2"] <- "blue" #"#6666FF"
levels(bcl_colors)[levels(bcl_colors) == "3"] <- "red" #"#FF6666"
bcl_colors <- as.vector(bcl_colors)
names(bcl_colors) <- names(bcl_com)

# pdf(file = paste0(draft_fig_path, "Fig_3_txn-clustering-union-ODs.pdf"),   # The directory you want to save the file in
#   width = 6, # The width of the plot in inches
#   height = 5) # The height of the plot in inches
par(mfrow=c(2,2), mar=rep(1,4))
# top row
# left: all slice spots tSNE embedding and labeled by cluster assignment
MERINGUE::plotEmbedding(emb, groups=bcl_com, colors = bcl_colors,
                     show.legend=FALSE, xlab=NA, ylab=NA,
                     verbose=TRUE)
# right: all slice spots tSNE embedding and labeled by tissuse slice assignment
MUDAN::plotEmbedding(emb, groups=bclCorpusUnion$slice,
                     show.legend=TRUE, xlab=NA, ylab=NA,
                     verbose=FALSE)
# bottom row
# left: spots labeld by cluster assignment and overlayed on each slice.
# Same clusters seem to be at similar spot locations on slices. Nice to see
MUDAN::plotEmbedding(bclCorpusUnion$pos, groups=bcl_com, colors = bcl_colors,
                     cex=1, xlab=NA, ylab=NA,
                     verbose=FALSE)
# right: spots colored by slice and slices overlayed
MUDAN::plotEmbedding(bclCorpusUnion$pos, groups=bclCorpusUnion$slice,
                     cex=1, xlab=NA, ylab=NA,
                     verbose=FALSE)
# dev.off()

# capture the txn cluster assignment of the spots
bclCorpusUnion$txn <- bcl_com

```

```{r}

bclUnionTxnClustCols <- c('1' = lighten('green', factor = 0.4),
                     '2' = lighten('blue', factor = 0.4),
                     '3' = lighten('red', factor = 0.4))

```

```{r}

# proxy theta for the txn clusters
bclProxyTheta_Union <- model.matrix(~ 0 + bclCorpusUnion$txn)
rownames(bclProxyTheta_Union) <- names(bclCorpusUnion$txn)
# fix names
colnames(bclProxyTheta_Union) <- c("1", "2", "3")
bclProxyTheta_Union <- as.data.frame.matrix(bclProxyTheta_Union)
dim(bclProxyTheta_Union)

```

```{r}

for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  # p <- bclCorpusUnion$pos[slice,]
  p <- posScaleAll[slice,]
  m <- as.matrix(bclProxyTheta_Union[slice,])

  plt <- vizAllTopics(theta = m,
                       pos = p,
                       topicOrder=colnames(m),
                       topicCols=as.vector(bclUnionTxnClustCols[colnames(m)]),
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 100, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s))
  print(plt)
  
}

```

# Compare Shared/Union Txn

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta),
                     m2 = as.matrix(bclProxyTheta_Union),
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

# LDA training

```{r}

# fit models to a range of K's that include 5 (5 original txn clusters) and
# 37 and 38 (for the cts in the scRNAseq refs; 38 originally but SL doesn't find and Meis2 so dropped)
ks <- seq(from = 2, to = 20, by = 1)

```

## Shared ODs

```{r}

# pdf(file = paste0(draft_fig_path, "Fig_3_bcl-1-4-shared-od-training.pdf"))
bclShared_LDAs <- fitLDA(counts = bclCorpus$slm, # 152 genes, 1029 spots
                        Ks = ks,
                        seed = 0,
                        ncores = 7,
                        plot = TRUE)
# dev.off()

```

### k=3 (histological)

```{r}

bcl_shared_k3 <- buildLDAobject(LDAmodel = optimalModel(bclShared_LDAs, opt = 3),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

### k=5 (kneed)

```{r}

bcl_shared_k5 <- buildLDAobject(LDAmodel = optimalModel(bclShared_LDAs, opt = "kneed"),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

### k=12 (min)

```{r}

bcl_shared_k12 <- buildLDAobject(LDAmodel = optimalModel(bclShared_LDAs, opt = "min"),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

## Union ODs

```{r}

# pdf(file = paste0(draft_fig_path, "Fig_3_bcl-1-4-union-od-training.pdf"))
bclUnion_LDAs <- fitLDA(counts = bclCorpusUnion$slm, # 372 genes 1029 spots
                        Ks = ks,
                        seed = 0,
                        ncores = 7,
                        plot = TRUE)
# dev.off()

```

### k=3 (histological)

```{r}

bcl_union_k3 <- buildLDAobject(LDAmodel = optimalModel(bclUnion_LDAs, opt = 3),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

### k=5 (kneed)

```{r}

bcl_union_k5 <- buildLDAobject(LDAmodel = optimalModel(bclUnion_LDAs, opt = "kneed"),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

### k=18 (min)

```{r}

bcl_union_k18 <- buildLDAobject(LDAmodel = optimalModel(bclUnion_LDAs, opt = "min"),
                      deepSplit = 4,
                      colorScheme = "rainbow")

```

# ---------------------------------

# adjusted positions for images

## he1

```{r}

he1 <- readJPEG('/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/HE_BrightfieldImages/HE_layer1_BC.jpg')

# alignment matrices came with these images
he_align1 <- matrix(unlist(read.table("/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/Alignments/Layer1_BC_transformation.txt")), nrow = 3, ncol = 3)

posScale1 <- bclCorpus$pos[bclCorpus$slice==1,]

# use the alignment file to adjust
posScale1[,"x"] <- posScale1[,"x"] * he_align1[1,1]
posScale1[,"y"] <- posScale1[,"y"] * he_align1[2,2]

# bring coordinates to positive values. set the x and y to start  at 0
posScale1[,"x"] <- posScale1[,"x"] - min(posScale1[,"x"])
posScale1[,"y"] <- posScale1[,"y"] - min(posScale1[,"y"])

# posScale1 <- MERINGUE::rotatePos(posScale1, theta = pi/(180/32))

# Flip the spots horizontally
posScale1[,"y"] <- posScale1[,"y"] + 2*((dim(he1)[1]/2) - posScale1[,"y"])
# posScale1[,"x"] <- posScale1[,"x"] + 2*((dim(he1)[2]/2) - posScale1[,"x"])

# posScale1 <- MERINGUE::rotatePos(posScale1, theta = pi/(180/30))

# set the x and y to start at 0 again, after flipping
posScale1[,"x"] <- posScale1[,"x"] - min(posScale1[,"x"])
posScale1[,"y"] <- posScale1[,"y"] - min(posScale1[,"y"])

posScale1 <- MERINGUE::rotatePos(posScale1, theta = pi/(180/91))

posScale1[,"x"] <- posScale1[,"x"] - min(posScale1[,"x"])
posScale1[,"y"] <- posScale1[,"y"] - min(posScale1[,"y"])

# final adjustment
# based on the Stahl et al spot coordinates (from Layer1_BC_count_matrix-1.tsv)
# that were adjusted with the alignment matrix, get the minimum x and y coords of these spots
# also double checked that the range of both sets of spot coordinates
# matched these, which they do. So just a matter of shifting spots in x and y direction
# equal to where the positions of the spots are in the stahl data
posScale1[,"x"] <- posScale1[,"x"] + 1500
posScale1[,"y"] <- posScale1[,"y"] + 1400

```

## he2

```{r}

he2 <- readJPEG('/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/HE_BrightfieldImages/HE_layer2_BC.jpg')

# alignment matrices came with these images
he_align2 <- matrix(unlist(read.table("/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/Alignments/Layer2_BC_transformation.txt")), nrow = 3, ncol = 3)

posScale2 <- bclCorpus$pos[bclCorpus$slice==2,]

# use the alignment file to adjust
posScale2[,"x"] <- posScale2[,"x"] * he_align2[1,1]
posScale2[,"y"] <- posScale2[,"y"] * he_align2[2,2]

# bring coordinates to positive values. set the x and y to start  at 0
posScale2[,"x"] <- posScale2[,"x"] - min(posScale2[,"x"])
posScale2[,"y"] <- posScale2[,"y"] - min(posScale2[,"y"])

# posScale1 <- MERINGUE::rotatePos(posScale1, theta = pi/(180/32))

# Flip the spots horizontally
posScale2[,"y"] <- posScale2[,"y"] + 2*((dim(he2)[1]/2) - posScale2[,"y"])
# posScale2[,"x"] <- posScale2[,"x"] + 2*((dim(he2)[2]/2) - posScale2[,"x"])

# set the x and y to start at 0 again, after flipping
posScale2[,"x"] <- posScale2[,"x"] - min(posScale2[,"x"])
posScale2[,"y"] <- posScale2[,"y"] - min(posScale2[,"y"])

posScale2 <- MERINGUE::rotatePos(posScale2, theta = pi/(180/78))

posScale2[,"x"] <- posScale2[,"x"] - min(posScale2[,"x"])
posScale2[,"y"] <- posScale2[,"y"] - min(posScale2[,"y"])

# final adjustment
# based on the Stahl et al spot coordinates (from Layer1_BC_count_matrix-1.tsv)
# that were adjusted with the alignment matrix, get the minimum x and y coords of these spots
# also double checked that the range of both sets of spot coordinates
# matched these, which they do. So just a matter of shifting spots in x and y direction
# equal to where the positions of the spots are in the stahl data
posScale2[,"x"] <- posScale2[,"x"] + 1200
posScale2[,"y"] <- posScale2[,"y"] + 2050

```

## he3

```{r}

he3 <- readJPEG('/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/HE_BrightfieldImages/HE_layer3_BC.jpg')

# alignment matrices came with these images
he_align3 <- matrix(unlist(read.table("/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/Alignments/Layer3_BC_transformation.txt")), nrow = 3, ncol = 3)

posScale3 <- bclCorpus$pos[bclCorpus$slice==3,]
# use the alignment file to adjust
posScale3[,"x"] <- posScale3[,"x"] * he_align3[1,1]
posScale3[,"y"] <- posScale3[,"y"] * he_align3[2,2]
# bring coordinates to positive values
posScale3[,"x"] <- posScale3[,"x"] - min(posScale3[,"x"])
posScale3[,"y"] <- posScale3[,"y"] - min(posScale3[,"y"])
# final adjustments

posScale3 <- MERINGUE::rotatePos(posScale3, theta = -pi/(180/32))
# for image 3, need to flip the spots across a vertical axis
# flip spots
posScale3[,"y"] <- posScale3[,"y"] + 2*((dim(he3)[1]/2) - posScale3[,"y"])
# posScale3[,"x"] <- posScale3[,"x"] + 2*((dim(he3)[2]/2) - posScale3[,"x"])
# then rotate the spots in radians
# posScale3 <- MERINGUE::rotatePos(posScale3, theta = -pi/(180/32))
# now reset the x and y to start at 0
posScale3[,"x"] <- posScale3[,"x"] - min(posScale3[,"x"])
posScale3[,"y"] <- posScale3[,"y"] - min(posScale3[,"y"])
# and finally shift by the start of the x and y coordinates for the stahl points (flipped across horizontal)
posScale3[,"x"] <- posScale3[,"x"] + 2300
posScale3[,"y"] <- posScale3[,"y"] + 2350

```

## he4

```{r}

he4 <- readJPEG('/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/HE_BrightfieldImages/HE_layer4_BC.jpg')

# alignment matrices came with these images
he_align4 <- matrix(unlist(read.table("/Users/brendan/Desktop/PostDoc/data/ImputeTranscriptomeFromHE/2016_Stahl/Alignments/Layer4_BC_transformation.txt")), nrow = 3, ncol = 3)

posScale4 <- bclCorpus$pos[bclCorpus$slice==4,]

# use the alignment file to adjust
posScale4[,"x"] <- posScale4[,"x"] * he_align4[1,1]
posScale4[,"y"] <- posScale4[,"y"] * he_align4[2,2]

# bring coordinates to positive values. set the x and y to start  at 0
posScale4[,"x"] <- posScale4[,"x"] - min(posScale4[,"x"])
posScale4[,"y"] <- posScale4[,"y"] - min(posScale4[,"y"])

# posScale1 <- MERINGUE::rotatePos(posScale1, theta = pi/(180/32))

# Flip the spots horizontally
posScale4[,"y"] <- posScale4[,"y"] + 2*((dim(he4)[1]/2) - posScale4[,"y"])
# posScale4[,"x"] <- posScale4[,"x"] + 2*((dim(he4)[2]/2) - posScale4[,"x"])

# set the x and y to start at 0 again, after flipping
posScale4[,"x"] <- posScale4[,"x"] - min(posScale4[,"x"])
posScale4[,"y"] <- posScale4[,"y"] - min(posScale4[,"y"])

posScale4 <- MERINGUE::rotatePos(posScale4, theta = pi/(180/5))

posScale4[,"x"] <- posScale4[,"x"] - min(posScale4[,"x"])
posScale4[,"y"] <- posScale4[,"y"] - min(posScale4[,"y"])

# final adjustment
# based on the Stahl et al spot coordinates (from Layer1_BC_count_matrix-1.tsv)
# that were adjusted with the alignment matrix, get the minimum x and y coords of these spots
# also double checked that the range of both sets of spot coordinates
# matched these, which they do. So just a matter of shifting spots in x and y direction
# equal to where the positions of the spots are in the stahl data
posScale4[,"x"] <- posScale4[,"x"] + 900
posScale4[,"y"] <- posScale4[,"y"] + 2300

```

combined adjusted positions

```{r}

posScaleAll <- rbind(posScale1,
      posScale2,
      posScale3,
      posScale4)

```

# ---------------------------------

# Shared ODs

## k=3

### Adjust proportions

```{r}

summary(bcl_shared_k3$theta)
ggplot(data = reshape2::melt(bcl_shared_k3$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_shared_k3_theta <- bcl_shared_k3$theta
bcl_shared_k3_theta[bcl_shared_k3_theta < 0.05] <- 0


# readjust proportions
bcl_shared_k3_theta <- bcl_shared_k3_theta/rowSums(bcl_shared_k3_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_shared_k3_theta[is.na(bcl_shared_k3_theta)] <- 0

summary(bcl_shared_k3_theta)

ggplot(data = reshape2::melt(bcl_shared_k3_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_shared_k3_theta <- bcl_shared_k3_theta[,which(colSums(bcl_shared_k3_theta) > 0)]

dim(bcl_shared_k3_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta),
                     m2 = bcl_shared_k3_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  # m <- bcl_shared_k3$theta[slice,]
  p <- bclCorpus$pos[slice,]
  # sliceTxnClusters <- bclCorpus$txn[slice]
  m <- as.matrix(bclProxyTheta[slice,])

  for(i in colnames(bcl_shared_k3_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_shared_k3_theta[slice,i]
    m_ <- cbind(m_, bcl_shared_k3_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}

# each topic on top of all the sections
for (i in seq(3)){

  m <- bcl_shared_k3_theta[,i]
  p <- bclCorpus$pos
  sliceTxnClusters <- bclCorpus$txn

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclTxnClustCols["3"], percent = 30))

  p <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 0.4,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(p)

}

# All topics for each section
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  m <- bcl_shared_k3_theta[slice,]
  p <- bclCorpus$pos[slice,]
  sliceTxnClusters <- bclCorpus$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

## k=5

### Adjust proportions

```{r}

summary(bcl_shared_k5$theta)
ggplot(data = reshape2::melt(bcl_shared_k5$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_shared_k5_theta <- bcl_shared_k5$theta
bcl_shared_k5_theta[bcl_shared_k5_theta < 0.05] <- 0


# readjust proportions
bcl_shared_k5_theta <- bcl_shared_k5_theta/rowSums(bcl_shared_k5_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_shared_k5_theta[is.na(bcl_shared_k5_theta)] <- 0

summary(bcl_shared_k5_theta)

ggplot(data = reshape2::melt(bcl_shared_k5_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_shared_k5_theta <- bcl_shared_k5_theta[,which(colSums(bcl_shared_k5_theta) > 0)]

dim(bcl_shared_k5_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta),
                     m2 = bcl_shared_k5_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  # m <- bcl_shared_k3$theta[slice,]
  p <- bclCorpus$pos[slice,]
  # sliceTxnClusters <- bclCorpus$txn[slice]
  m <- as.matrix(bclProxyTheta[slice,])

  for(i in colnames(bcl_shared_k5_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_shared_k5_theta[slice,i]
    m_ <- cbind(m_, bcl_shared_k5_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}


# each topic on top of all the sections
for (i in seq(5)){

  m <- bcl_shared_k5_theta[,i]
  p <- bclCorpus$pos
  sliceTxnClusters <- bclCorpus$txn

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclTxnClustCols["3"], percent = 30))

  p <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 0.4,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(p)

}

# All topics for each section
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  m <- bcl_shared_k5_theta[slice,]
  p <- bclCorpus$pos[slice,]
  sliceTxnClusters <- bclCorpus$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

## k=12

### Adjust proportions

```{r}

summary(bcl_shared_k12$theta)
ggplot(data = reshape2::melt(bcl_shared_k12$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_shared_k12_theta <- bcl_shared_k12$theta
bcl_shared_k12_theta[bcl_shared_k12_theta < 0.05] <- 0


# readjust proportions
bcl_shared_k12_theta <- bcl_shared_k12_theta/rowSums(bcl_shared_k12_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_shared_k12_theta[is.na(bcl_shared_k12_theta)] <- 0

summary(bcl_shared_k12_theta)

ggplot(data = reshape2::melt(bcl_shared_k12_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_shared_k12_theta <- bcl_shared_k12_theta[,which(colSums(bcl_shared_k12_theta) > 0)]

dim(bcl_shared_k12_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta),
                     m2 = bcl_shared_k12_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  # m <- bcl_shared_k3$theta[slice,]
  p <- bclCorpus$pos[slice,]
  # sliceTxnClusters <- bclCorpus$txn[slice]
  m <- as.matrix(bclProxyTheta[slice,])

  for(i in colnames(bcl_shared_k12_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_shared_k12_theta[slice,i]
    m_ <- cbind(m_, bcl_shared_k12_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}


# each topic on top of all the sections
for (i in seq(12)){

  m <- bcl_shared_k12_theta[,i]
  p <- bclCorpus$pos
  sliceTxnClusters <- bclCorpus$txn

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclTxnClustCols["3"], percent = 30))

  p <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 0.4,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(p)

}

# All topics for each section
for (s in unique(bclCorpus$slice)){
  slice <- bclCorpus$slice == s
  m <- bcl_shared_k12_theta[slice,]
  p <- bclCorpus$pos[slice,]
  sliceTxnClusters <- bclCorpus$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

# Union ODs

## k=3

### Adjust proportions

```{r}

summary(bcl_union_k3$theta)
ggplot(data = reshape2::melt(bcl_union_k3$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_union_k3_theta <- bcl_union_k3$theta
bcl_union_k3_theta[bcl_union_k3_theta < 0.05] <- 0


# readjust proportions
bcl_union_k3_theta <- bcl_union_k3_theta/rowSums(bcl_union_k3_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_union_k3_theta[is.na(bcl_union_k3_theta)] <- 0

summary(bcl_union_k3_theta)

ggplot(data = reshape2::melt(bcl_union_k3_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_union_k3_theta <- bcl_union_k3_theta[,which(colSums(bcl_union_k3_theta) > 0)]

dim(bcl_union_k3_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta_Union),
                     m2 = bcl_union_k3_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  p <- bclCorpusUnion$pos[slice,]
  m <- as.matrix(bclProxyTheta_Union[slice,])

  for(i in colnames(bcl_union_k3_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_union_k3_theta[slice,i]
    m_ <- cbind(m_, bcl_union_k3_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclUnionTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}

# each topic on top of all the sections
for (i in seq(3)){

  m <- bcl_union_k3_theta[,i]
  p <- bclCorpusUnion$pos
  sliceTxnClusters <- bclCorpusUnion$txn

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclUnionTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclUnionTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclUnionTxnClustCols["3"], percent = 30))

  p <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 0.4,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(p)

}

# All topics for each section
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  m <- bcl_union_k3_theta[slice,]
  p <- bclCorpusUnion$pos[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

## k=5

### Adjust proportions

```{r}

summary(bcl_union_k5$theta)
ggplot(data = reshape2::melt(bcl_union_k5$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_union_k5_theta <- bcl_union_k5$theta
bcl_union_k5_theta[bcl_union_k5_theta < 0.05] <- 0


# readjust proportions
bcl_union_k5_theta <- bcl_union_k5_theta/rowSums(bcl_union_k5_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_union_k5_theta[is.na(bcl_union_k5_theta)] <- 0

summary(bcl_union_k5_theta)

ggplot(data = reshape2::melt(bcl_union_k5_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_union_k5_theta <- bcl_union_k5_theta[,which(colSums(bcl_union_k5_theta) > 0)]

dim(bcl_union_k5_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta_Union),
                     m2 = bcl_union_k5_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  p <- bclCorpusUnion$pos[slice,]
  m <- as.matrix(bclProxyTheta_Union[slice,])

  for(i in colnames(bcl_union_k5_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_union_k5_theta[slice,i]
    m_ <- cbind(m_, bcl_union_k5_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclUnionTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 0.4, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}

# each topic on top of all the sections
for (i in seq(5)){

  m <- bcl_union_k5_theta[,i]
  p <- bclCorpusUnion$pos
  sliceTxnClusters <- bclCorpusUnion$txn

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclUnionTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclUnionTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclUnionTxnClustCols["3"], percent = 30))

  p <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 0.4,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(p)

}

# All topics for each section
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  m <- bcl_union_k5_theta[slice,]
  p <- bclCorpusUnion$pos[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

## ** k=18

### Adjust proportions

```{r}

summary(bcl_union_k18$theta)
ggplot(data = reshape2::melt(bcl_union_k18$theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


bcl_union_k18_theta <- bcl_union_k18$theta
bcl_union_k18_theta[bcl_union_k18_theta < 0.05] <- 0


# readjust proportions
bcl_union_k18_theta <- bcl_union_k18_theta/rowSums(bcl_union_k18_theta)
# if NAs because all cts are 0 in a spot, replace with 0
bcl_union_k18_theta[is.na(bcl_union_k18_theta)] <- 0

summary(bcl_union_k18_theta)

ggplot(data = reshape2::melt(bcl_union_k18_theta)) +
  geom_jitter(aes(x = reorder(Var2, -value), y = value), position=position_jitter(0.1), cex=0.1) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        # panel.grid = element_blank(),
        # axis.line=element_blank(),
        # axis.text.x=element_blank(),
        # axis.text.y=element_blank(),
        # axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank()
        ) +
  ggtitle("")


# drop ant cts that are 0 for all pixels
bcl_union_k18_theta <- bcl_union_k18_theta[,which(colSums(bcl_union_k18_theta) > 0)]

dim(bcl_union_k18_theta)

```

Because LDA is looking for gene occurrences in pixels, and pixels share topics at different mixtures. Whereas the txn clustering groups pixels based on their overall txn profile. Because this is a lot more heterogeneous than the MOB, I wouldn't expect much correlation between topics and txn clusters.

### Txn cluster compare

```{r}

corMtx <- getCorrMtx(m1 = as.matrix(bclProxyTheta_Union),
                     m2 = bcl_union_k18_theta,
                     type = "t")

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor.pdf"))
gplots::heatmap.2(x = corMtx,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

# pdf(file = paste0(fig_path, "Fig1_B-4_FN7_k8_mjrcts_thetaCor_pairs.pdf"))
# pair up best matching cell class to a predicted topic
pairs <- lsatPairs(corMtx)
corMtx_paired <- corMtx[pairs$rowix, pairs$colsix]

gplots::heatmap.2(x = corMtx_paired,
                  Rowv = NA, # to order topics based on pairs and not dendrogram
                  Colv = NA,
                  density.info = "none",
                  trace = "none",
                  col = correlation_palette,
                  breaks = correlation_breaks,
                  key.xlab = "Correlation",
                  # xlab = "Major cell class",
                  # ylab = "Predicted topics",
                  key.title = NA,
                  cexRow = 1,
                  cexCol = 0.9,
                  # lhei = c(1,3),
                  margins = c(7,4))
# dev.off()

diags <- diag(corMtx_paired)

# hist(rctd_vs_lda_theta, breaks = 10)
summary(diags)

# pdf(file = paste0(fig_path, "Fig_S6_SL-mob-vs-cortex-corr-boxplot.pdf"),   # The directory you want to save the file in
    # width = 5, # The width of the plot in inches
    # height = 3) # The height of the plot in inches
par( mfrow = c(1,3) )
boxplot(diags,
        medcol = "black", # medium line color
        medlty = 1, # line type
        medlwd = 1, # line width
        boxfill = "white", # color fill of box
        boxcol = "black", # outline color of box
        boxlty = 1, # box line type
        boxlwd = 1, # width of box line
        whisklty = 1, # whisker type
        whisklwd = 1,
        staplelwd = 0, # the ends of the whiskers
        outpch = 1, # type of outlies
        outcex = 0,
        main = "",
        cex.main = 1,
        xlab = "",
        ylab = "Correlation") # size of outliers
stripchart(diags, vertical = TRUE,
    method = "jitter", add = TRUE, pch = 20, col = 'black')
# dev.off()

```

not as strong correlation but honestly still pretty good. Given the heterogeneousness of the tissue.
Perhaps worth stating.

### viz

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  # p <- bclCorpusUnion$pos[slice,]
  p <- posScaleAll[slice,]
  m <- as.matrix(bclProxyTheta_Union[slice,])

  for(i in colnames(bcl_union_k18_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_union_k18_theta[slice,i]
    m_ <- cbind(m_, bcl_union_k18_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclUnionTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 100, # different size piecharts for mOB data sets
                       lwd = 0.3,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

```{r}

# each topic on top of all the sections
for (i in seq(18)){
  slice <- bclCorpusUnion$slice == 1
  m <- bcl_union_k18_theta[slice,i]
  # p <- bclCorpusUnion$pos[slice,]
  p <- posScaleAll[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("black", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclUnionTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclUnionTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclUnionTxnClustCols["3"], percent = 30))

  plt <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = sliceTxnClusters,
                 group_cols = group_cc,
                 r = 100,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(plt)

}

# All topics for each section
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  m <- bcl_union_k18_theta[slice,]
  # p <- bclCorpusUnion$pos[slice,]
  p <- posScaleAll[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]
  
  plt <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=gg_color_hue(ncol(m)),
               groups = sliceTxnClusters,
               group_cols = c("black", "black", "black"),
               r = 100, # different size piecharts for mOB data sets
               lwd = 0.3,
               showLegend = TRUE,
               plotTitle = NA)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  plt <- plt + guides(fill=guide_legend(ncol=2))
  print(plt)
}

```

```{r}

# Each topic versus the 3 txn clsuters, each section separately
for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  # p <- bclCorpusUnion$pos[slice,]
  p <- posScaleAll[slice,]
  m <- as.matrix(bclProxyTheta_Union[slice,])

  for(i in colnames(bcl_union_k18_theta)){
    # for each topic, subtract its proportion from the txn cluster.
    # Adjust the 0's that went negative back to 0
    m_ <- m - bcl_union_k18_theta[slice,i]
    m_ <- cbind(m_, bcl_union_k18_theta[slice,i])
    colnames(m_) <- c(colnames(m), paste0("Ct.",i))
    m_[m_ < 0] <- 0
    
    cc <- as.vector(bclUnionTxnClustCols[colnames(m)])
    cc <- c(cc, "black")
    
    plt <- vizAllTopics(theta = m_,
                       pos = p,
                       topicOrder=colnames(m_),
                       topicCols=cc,
                       groups = rep("0", dim(m)[1]),
                       group_cols = c("0" = "black"),
                       r = 100, # different size piecharts for mOB data sets
                       lwd = 0.1,
                       showLegend = TRUE,
                       plotTitle = paste("Txn clusters section", s, "CT", i))
    print(plt)
  }
  
}

```

# Convert RGB to grayscale for intensity measurements

```{r rgb-to-pixel-intensity, fig.width=10, fig.height=11, eval=FALSE}

# https://en.wikipedia.org/wiki/Grayscale

# convert RGB values from gamma-compressed to linear scale
c.linear <- (function(c) if (c <= 0.04045) {
  c/12.92 } else {
    ((c + 0.055)/1.055)^2.4 } )

tiffLinear <- structure(sapply(he1, c.linear), dim=dim(he1))

# take weighted sum to calculate linear Luminance
tiffYlinear <- (tiffLinear[, , 1] * 0.2126) + (tiffLinear[, , 2] * 0.7152) + (tiffLinear[, , 3] * 0.0722)

# transform back to gamma-compressed scale
srgb <- (function(y) if (y <= 0.0031308) {12.92*y} else { (1.055*(y^(1/2.2))) - 0.055 })

tiffGray <- structure(sapply(tiffYlinear, srgb), dim=dim(tiffYlinear))

# library(wvtool)
# tiffGray <- rgb2gray(tiff)

# check if it worked
plot.new()
plot(c(0,dim(tiffGray)[2]), c(0,dim(tiffGray)[1]))
rasterImage(tiffGray[,],
            0, 0,
            dim(tiffGray)[2], dim(tiffGray)[1])

```

```{r}

for (i in c(11, 13, 17)){
  slice <- bclCorpusUnion$slice == 1
  m <- bcl_union_k18_theta[slice,i]
  p <- posScaleAll[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]

  other <- 1 - m
  m <- cbind(m, other)
  colnames(m) <- c(as.character(i), "other")

  cc <- c(
    transparentCol("magenta", percent = 30),
    # transparentCol("blue", percent = 50),
    transparentCol("white", percent = 100))

  group_cc <- c("1" = transparentCol(bclUnionTxnClustCols["1"], percent = 30),
                "2" = transparentCol(bclUnionTxnClustCols["2"], percent = 30),
                "3" = transparentCol(bclUnionTxnClustCols["3"], percent = 30))

  plt <- vizAllTopics(theta = m,
                 pos = p,
                 topicOrder=seq(ncol(m)),
                 topicCols=cc,
                 groups = NA,
                 group_cols = NA,
                 r = 100,
                 lwd = 0.1,
                 showLegend = TRUE,
                 plotTitle = NA,
                 overlay = tiffGray)
  # ggsave(filename = paste0("bcl_k5-layers-topic-", i, ".pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)

  print(plt)

}

```


```{r}

green_ramp <- colorRampPalette(c(lighten('green', factor = 0.1), darken('green', factor = 2)))
clust_1 <- green_ramp(9)

blue_ramp <- colorRampPalette(c(lighten('blue', factor = 0.1), darken('blue', factor = 2)))
clust_2 <- blue_ramp(5)

red_ramp <- colorRampPalette(c(lighten('red', factor = 0.1), darken('red', factor = 2)))
clust_3 <- red_ramp(4)

perc_transp <- 40
bcl_union_k18_cols <- c("1" = transparentCol(clust_2[1], percent = perc_transp),
                       "2" = transparentCol(clust_2[2], percent = perc_transp),
                       "3" = transparentCol(clust_1[1], percent = perc_transp),
                       "4" = transparentCol(clust_1[2], percent = perc_transp),
                       "5" = transparentCol(clust_2[3], percent = perc_transp),
                       "6" = transparentCol(clust_2[4], percent = perc_transp),
                       "7" = transparentCol(clust_1[3], percent = perc_transp),
                       "8" = transparentCol(clust_1[4], percent = perc_transp),
                       "9" = transparentCol(clust_1[5], percent = perc_transp),
                       "10" = transparentCol(clust_1[6], percent = perc_transp),
                       "11" = transparentCol(clust_3[1], percent = perc_transp),
                       "12" = transparentCol(clust_3[2], percent = perc_transp),
                       "13" = transparentCol(clust_2[5], percent = perc_transp),
                       "14" = transparentCol(clust_3[3], percent = perc_transp),
                       "15" = transparentCol(clust_1[7], percent = perc_transp),
                       "16" = transparentCol(clust_1[8], percent = perc_transp),
                       "17" = transparentCol(clust_3[4], percent = perc_transp),
                       "18" = transparentCol(clust_1[9], percent = perc_transp)
                       )


m <- bcl_union_k18_theta
p <- bclCorpusUnion$pos
plt <- vizAllTopics(theta = m,
                     pos = p,
                     topicOrder=seq(ncol(m)),
                     topicCols=as.vector(bcl_union_k18_cols[as.character(seq(18))]),
                     groups = rep("0", dim(m)[1]),
                     group_cols = c("0" = "black"),
                     r = 0.4, # different size piecharts for mOB data sets
                     lwd = 0.1,
                     showLegend = TRUE,
                     plotTitle = "")
print(plt)


for (s in unique(bclCorpusUnion$slice)){
  slice <- bclCorpusUnion$slice == s
  m <- bcl_union_k18_theta[slice,]
  p <- bclCorpusUnion$pos[slice,]
  sliceTxnClusters <- bclCorpusUnion$txn[slice]
  
  p <- vizAllTopics(theta = m,
               pos = p,
               topicOrder=seq(ncol(m)),
               topicCols=as.vector(bcl_union_k18_cols[as.character(seq(18))]),
               groups = rep("0", dim(m)[1]),
               group_cols = c("0" = "black"),
               r = 0.4, # different size piecharts for mOB data sets
               lwd = 0.1,
               showLegend = TRUE,
               plotTitle = s)
  # ggsave(filename = paste0("bcl_k5-layer-", s, "-all-topics.pdf"),
  #        device = "pdf",
  #        path = draft_fig_path,
  #        scale = 1.5,
  #        width = 5,
  #        height = 4,
  #        units = c("in"),
  #        dpi = 600)
  
  print(p)
}

```

```{r}

for (s in unique(bclCorpusUnion$slice)[1]){
  slice <- bclCorpusUnion$slice == s
  m <- bcl_union_k18_theta[slice,]
  p <- bclCorpusUnion$pos[slice,]
  
  cc <- as.factor(bcl_union_k18_cols)
  names(cc) <- colnames(m)
  
  vizTopicClusters(theta = m,
                   pos = p,
                   clusters = as.factor(bcl_union_k18_cols),
                   sharedCol = TRUE,
                   groups = rep("0", dim(m)[1]),
                   group_cols = c("0" = "black"),
                   r = 0.4,
                   lwd = 0.3,
                   showLegend = TRUE,
                   plotTitle = "slice 1")
}

```

# ---------------------------------
# GO

```{r}

library(liger)
data(org.Hs.GO2Symbol.list) ## load built in GO gene sets

# list of GO terms and associated genes
go.env <- org.Hs.GO2Symbol.list

library(GO.db)
library(AnnotationDbi)

# annotate the GO terms
# dataframe of GO ID and the associated term annotation
desc <- AnnotationDbi::select(
  GO.db,
  keys = names(go.env),
  columns = c("TERM"),
  multiVals = 'CharacterList'
)

# append GO term list names with their annotations
names(go.env) <- paste(names(go.env), desc$TERM)

```

## Shared

```{r}

# list of gene sets
# go.env

# the 152 genes
corpusGenes <- colnames(bclCorpus$corpus)

# filter down list to only include sets that overlap the corpus genes


includeSets <- lapply(seq(length(go.env)), function(l){
  setGenes <- go.env[l][[1]]
  if(length(intersect(corpusGenes, setGenes)) >= 1 ){  # round(length(corpusGenes)*0.05)
    l
  }
})
# unlist(includeSets)

filtGOsets_shared <- go.env[unlist(includeSets)]
print(length(filtGOsets_shared))

```

### k=3

```{r}

gset <- filtGOsets_shared
for (i in seq(3)){
  
  vals <- sort(bcl_shared_k3$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 10))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

### k=5

```{r}

gset <- filtGOsets_shared
for (i in seq(5)){
  
  vals <- sort(bcl_shared_k5$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 10))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

### k=12

```{r}

gset <- filtGOsets_shared
for (i in seq(12)){
  
  vals <- sort(bcl_shared_k12$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 10))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

## Union

```{r}

# list of gene sets
# go.env

# the 372 genes
corpusGenes <- colnames(bclCorpusUnion$corpus)

# filter down list to only include sets that overlap the corpus genes

includeSets <- lapply(seq(length(go.env)), function(l){
  setGenes <- go.env[l][[1]]
  if(length(intersect(corpusGenes, setGenes)) >= 1 ){  # round(length(corpusGenes)*0.05)
    l
  }
})
# unlist(includeSets)

filtGOsets_union <- go.env[unlist(includeSets)]
print(length(filtGOsets_union))

```

### k=3

```{r}

gset <- filtGOsets_union
for (i in seq(3)){
  
  vals <- sort(bcl_union_k3$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 10))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

### k=5

```{r}

gset <- filtGOsets_union
for (i in seq(5)){
  
  vals <- sort(bcl_union_k5$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 30))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

### ** k=18

```{r}

gset <- filtGOsets_union
for (i in seq(18)){
  
  vals <- sort(bcl_union_k18$beta[i,], decreasing=TRUE)
  print(head(vals))
  gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
  print(head(gsea.results[order(gsea.results$p.val),], n = 30))
  
  # filter for top hits
  gsea.sig <- gsea.results[gsea.results$q.val < 0.1,]
  gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
  print(dim(gsea.sig))
  print(head(gsea.sig, n = 30))
  
  # visualize the top 5
  # for (j in seq(3)){
  #   pdf(file = paste0(gsea_bcl_draft_figs, "Fig_3_gsea-GO-bcl_k5-topic-", i, "-", rownames(gsea.sig)[j], ".pdf"),
  #     width = 6, height = 5)
  #   gsea(values=vals, geneset=gset[[rownames(gsea.sig)[j]]], plot=TRUE, rank=TRUE, main = paste("topic", i, rownames(gsea.sig)[j]))
  #   dev.off()
  # }
}

```

17 has positive score immune terms

```{r}

gset <- filtGOsets_union
vals <- sort(bcl_union_k18$beta[17,], decreasing=TRUE)
print(head(vals))
gsea.results <- iterative.bulk.gsea(values=vals, set.list=gset, rank=TRUE)
print(head(gsea.results[order(gsea.results$p.val),], n = 30))

# filter for top hits
gsea.sig <- gsea.results[gsea.results$q.val < 0.05,]
gsea.sig <- gsea.sig[order(gsea.sig$p.val),]
print(dim(gsea.sig))
print(head(gsea.sig, n = 30))

```

```{r}

gsea.sig$term <- rownames(gsea.sig)

highlights <- ifelse(grepl("immune", arrange(gsea.sig, -q.val)$term), "red", "black")


ggplot(data = gsea.sig, aes(x = reorder(term, -log10(q.val)), y = -log10(q.val))) +
  geom_col() +
  coord_flip()  +
  theme(axis.text.y = element_text(colour = highlights)) +
  geom_hline(yintercept=-log10(0.05), color = "red", linetype = "dashed")

```

### viz gexp

of signature and also top genes

CD74 top by a lot
It is involved in the formation and transport of MHC class II peptide complexes for the generation of CD4+ T cell responses.

CXCL10 is secreted by several cell types in response to IFN-. These cell types include monocytes, endothelial cells and fibroblasts.


way to viz each section separately but also way to overlay the gexp for all the sections?

```{r}

correlation_palette2 <- colorRampPalette(c("blue", "white", "red"))(n = 209)
correlation_breaks2 = c(seq(-1,-0.01,length=100),
                       seq(-0.009,0.009,length=10),
                       seq(0.01,10,length=100))

vizGeneCounts2 <- function(df, gene,
                          groups = NA,
                          group_cols = NA,
                          size = 3, stroke = 0.1, alpha = 1,
                          plotTitle = NA,
                          showLegend = TRUE) {

  counts <- df[,gene]

  # color spots by group:
  if (is.na(groups[1]) == TRUE) {
    groups <- " "
    # stroke <- 0.5
  } else {
    groups <- as.character(groups)
  }
  if (is.na(group_cols[1]) == TRUE) {
    group_cols <- c(" " = "white")
  }

  p <- ggplot() +
    geom_point(data = df, aes(x=x, y=y, fill=counts, color = groups),
               shape = 21,
               stroke = stroke, size = size,
               alpha = alpha) +
    scale_fill_gradientn(colours = c("white", "red")) +
    scale_color_manual(values = group_cols)
  
  if (showLegend == FALSE) {
    p <- p + guides(fill=FALSE)
  }
  if (is.na(plotTitle) == FALSE) {
    p <- p + ggtitle(plotTitle)
  }
  
  p <- p +
    theme(
      #panel.background = element_rect(fill = "white"),
      panel.grid = element_blank(),
      axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      panel.background=element_blank())
    # theme_classic()
  print(p)
}

```


```{r}

slice <- bclCorpusUnion$slice == 1

# pos <- bclCorpusUnion$pos[slice,]
pos <- posScaleAll[slice,]
pixels <- rownames(pos)[slice]

# top genes from topic gene probabilities
vals <- sort(bcl_union_k18$beta[17,], decreasing=TRUE)[1:3]

bcl_genes <- rownames(bcl_CPM[unionBCLGenes, pixels])
setGenes <- filtGOsets_union[["GO:0045089 positive regulation of innate immune response"]]
sig_genes <- intersect(bcl_genes, setGenes)
sig_genes

sig_gexp <- as.matrix(scale(colSums(bcl_CPM[sig_genes, pixels]))) #[,1]
colnames(sig_gexp) <- c("GO:0045089")

topic_gene_counts <- t(bcl_CPM[labels(vals), pixels])

gexp <- cbind(sig_gexp, topic_gene_counts)


df <- merge(pos, as.data.frame(gexp), by=0)
rownames(df) <- df$Row.names
df

# df <- t(bcl_CPM)[pixels, genes]
# df <- merge(pos, as.data.frame(df), by=0)
# rownames(df) <- df$Row.names
# df

```

```{r, fig.height=5, fig.width=6}

for(g in colnames(df)[4:7]){
  vizGeneCounts(df = df,
                  gene = g,
                  groups = NA,
                  group_cols = "black",
                  size = 5, stroke = 0.5,
                  alpha = 0.5,
                  plotTitle = g,
                  showLegend = TRUE)
}

```

```{r}

vals <- sort(bcl_union_k18$beta[17,], decreasing=TRUE)
dat <- data.frame(values = as.vector(vals), genes = names(vals), order = seq(length(vals)))

# Hide all of the text labels.
dat$selectedLabels <- ""
# Let's just label these items.
ix_label <- seq(5)
dat$selectedLabels[ix_label] <- dat$genes[ix_label]

ggplot(data = dat) +
  geom_col(aes(x = order, y = values)) +
  labs(title = "Cell-type 17",
       x = "Genes", y = "Normalized expression") +
  # coord_cartesian(clip = "off") +
  geom_text_repel(data = dat, mapping = aes(x = order, y = values, label = selectedLabels), size = 5, 
                  min.segment.length = 0.01, seed = 42, box.padding = 0.1, max.overlaps = Inf, point.padding = 0.01,
                  # fill = "white", xlim = c(-Inf, Inf), ylim = c(-Inf, Inf), 
                  nudge_x = 10, direction = "y", hjust = "left",
                  # segment.curvature = -0.1, segment.ncp = 3, segment.angle = 20)
  ) +
  theme_classic()

```


# fgsea

# -----------------------------------

Try the other msgea categories? immune?


```{r}

fitLDA2 <- function(counts, Ks = seq(2, 10, by = 2), seed = 0, testSize = NULL,
                   ncores = parallel::detectCores(logical = TRUE) - 1,
                   plot = TRUE) {
  
  if (is.null(testSize)){
    set.seed(seed)
    testingPixels <- seq(nrow(counts))
    fittingPixels <- seq(nrow(counts))
  } else if ((0 < testSize) & (testSize < 1.0)){
    cat("Splitting pixels into", testSize*100, "% and", 100-testSize*100, "% testing and fitting corpuses", "\n")
    set.seed(seed)
    testingPixels <- sample(nrow(counts), round(nrow(counts)*testSize))
    fittingPixels <- seq(nrow(counts))[-testingPixels]
  } else {
    stop("`testSize` must be NULL or decimal between 0 and 1")
  }
  
  ## counts must be pixels (rows) x genes (cols) matrix
  corpusFit <- slam::as.simple_triplet_matrix((as.matrix(counts[fittingPixels,])))
  corpusTest <- slam::as.simple_triplet_matrix((as.matrix(counts[testingPixels,])))
  
  # if (slam::is.simple_triplet_matrix(counts) == TRUE){
  #   corpus <- counts
  # } else {
  #   corpus <- slam::as.simple_triplet_matrix(t(as.matrix(counts)))
  # }
  
  controls <- list(seed = seed,
                   verbose = 1, keep = 1, estimate.alpha = TRUE)
  
  start_time <- Sys.time()
  fitted_models <- parallel::mclapply(Ks, function(k) {
    topicmodels::LDA(corpusFit, k=k, control = controls)
  },
  mc.cores = ncores
  )
  names(fitted_models) <- Ks
  total_t <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  print(sprintf("Time to fit LDA models was %smins", total_t))
  
  print("Computing perplexity for each fitted model...")
  pScores <- unlist(lapply(fitted_models, function(model){
    p <- topicmodels::perplexity(model, newdata = corpusTest)
    return(p)
  }))

  ## Kneed algorithm
  kOpt1 <- Ks[where.is.knee(pScores)]
  ## Min
  kOpt2 <- Ks[which(pScores == min(pScores))]
  
  ## check number of predicted cell-types at low proportions
  out <- lapply(1:length(Ks), function(i) {
    apply(getBetaTheta(fitted_models[[i]])$theta, 2, mean)
    })
  ## number of topics present at fewer than 5% on average across pixels
  numrare <- unlist(lapply(out, function(x) sum(x < 0.05)))
  kOpt3 <- Ks[where.is.knee(numrare)]

  if(plot) {
    plot(Ks, pScores, ylab = "perplexity", xlab = "K", main = "K vs perplexity")
    abline(v = kOpt1, col='blue')
    abline(v = kOpt2, col='red')
    legend(x = "top",
           legend = c("kneed", "min"), col = c("blue", "red"), lty = 1, lwd = 1)
    
    plot(Ks, numrare, ylab = "number of cell-types at < 5% mean proportion", xlab = "K", main = "K vs number of rare predicted cell-types")
    abline(v = kOpt3, col='red')
    legend(x = "top",
           legend = c("kneed"), col = c("red"), lty = 1, lwd = 1)
  }

  return(list(models = fitted_models,
              kOpt1 = kOpt1,
              kOpt2 = kOpt2,
              kOpt3 = kOpt3,
              numRare = numrare,
              perplexities = pScores,
              fitCorpus = corpusFit,
              testCorpus = corpusTest))
}

```


```{r}

ks <- seq(from = 2, to = 30, by = 2)

```


```{r}

# pdf(file = paste0(draft_fig_path, "Fig_3_bcl-1-4-union-od-training.pdf"))
bclUnion_LDAs2 <- fitLDA2(counts = bclCorpusUnion$corpus, # 372 genes 1029 spots
                        Ks = ks,
                        testSize = NULL,
                        seed = 0,
                        ncores = 7,
                        plot = TRUE)
# dev.off()

# pdf(file = paste0(draft_fig_path, "Fig_3_bcl-1-4-union-od-training.pdf"))
bclUnion_LDAs3 <- fitLDA2(counts = bclCorpusUnion$corpus, # 372 genes 1029 spots
                        Ks = ks,
                        testSize = 0.2,
                        seed = 0,
                        ncores = 7,
                        plot = TRUE)
# dev.off()

```

```{r}

ks <- seq(from = 2, to = 20, by = 1)
## look at distribution of cell-type proportions
out <- lapply(1:length(ks), function(i) {
  apply(getBetaTheta(bclUnion_LDAs$models[[i]])$theta, 2, mean)
})
## number of topics present at fewer than 5% on average across pixels
numrare <- unlist(lapply(out, function(x) sum(x < 0.05)))
ks[where.is.knee(numrare)]
plot(ks, numrare, type="l")

```












